---
title: "7dpf naglu"
author: "LR"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Library


```{r library, include=FALSE}
# data manipulation
library(tidyverse)
library(magrittr)
library(readxl)

# statistical analysis
library(lme4)
library(broom)
library(car)
library(performance)
library(emmeans)
library(MASS)

# data visualstion
library(scales)
library(ggpubr)
library(ggeasy)
library(ggfortify)
library(ggbeeswarm)
library(kableExtra)
library(ggplate)

theme_set(theme_bw())
```




```{r joining data}


fulldata <-
  read_excel("final data after fixing sensitivity issue.xlsx") %>%
    rename(
    "Trial" = `...2`,
    "Position" = `...3`,
    "Bin" = `...4`,
    "Distance_travelled" = "Distance moved center-point Total mm",
    "Freq_in_centre" = "In zone centre / center-point Frequency" ,
    "Time_in_centre" = "In zone centre / center-point Cumulative Duration s",
    "cum_movement" = "Movement Moving / center-point Cumulative Duration s"
    ) %>%
  dplyr::select(-c("Independent Variable Independent Variable Treatment", `...1`)) %>%
    mutate(Bin = factor(Bin, levels = c("Start-0:20:00", "0:20:00-0:40:00", "0:40:00-1:00:00")),
       Trial = str_remove(Trial, pattern = "Trial     ") %>%
         as.factor()) 

  
  meta <-
  read_excel("Genotype 7dpf only 49-120.xlsx") %>%
  mutate(Genotype = factor(Genotype, levels = c('wt', 'het', 'hom')),
         Trial = as.factor(Trial),
         Position = as.factor(Position)
  )

fulldata <-
  meta %>%
  left_join(fulldata, by = c("Trial", "Position")) %>%
  na.omit


#The daniovision calibration length was set at 1000 mm instead of 120. correcting it by multiplying by 0.12
fulldata$Distance_travelled <- fulldata$Distance_travelled * 0.12







```

```{r proportions}



 totaldist <-
  fulldata %>%
  group_by(fish_id) %>%
  mutate(Total_distance = sum(Distance_travelled)) %>%
  dplyr::select(fish_id, Trial, Position, Genotype, Total_distance, Freq_in_centre) %>%
  dplyr::distinct(Total_distance, .keep_all = TRUE) %>%
  ungroup() %>%
  droplevels()




geno.props.plot <- function(xaxis, input) {
  ggplot(data = input, aes(x = xaxis, fill = Genotype)) +
    geom_bar(mapping = aes(x = Genotype)) +
    scale_fill_viridis_d(end = 0.85)
}


geno.props.plot(input = totaldist, xaxis = "Genotype")


```


```{r exporting data}

totaldist %>% saveRDS("B.7.totaldist.rds")

fulldata %>% saveRDS("B.7.fulldata.rds")

```





Plotting overall distance travelled by genotype

```{r overall distance plot}

totaldist %>%
 ggplot(aes(x = Genotype, y = Total_distance)) +
    geom_jitter(aes(colour = Genotype,
                  shape = Trial),
              size = 4) +
  geom_boxplot(aes(fill = Genotype),
               alpha = 0.25,
               outlier.shape = NA
  ) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
    scale_y_continuous(labels = comma, name = "Distance (mm)") +
  theme(text = element_text(size = 15))
#+
  #ggtitle("Total distance travelled in an hour (mm)")
  
ggsave(filename = "Total distance in an hour.png", width = 15, height = 10, units = "cm", dpi = 300, scale = 1.5)

  
  
fulldata %>%
 ggplot(aes(x = Genotype, y = Distance_travelled)) +
    geom_jitter(aes(colour = Genotype,
                  shape = Trial),
              size = 4) +
  geom_boxplot(aes(fill = Genotype),
               alpha = 0.25,
               outlier.shape = NA
  ) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
    scale_y_continuous(labels = comma, name = "Distance (mm)") +
  facet_wrap(~Bin) +
  theme(text = element_text(size = 15))

ggsave(filename = "Distance per bin.png", width = 20, height = 10, units = "cm", dpi = 300, scale = 1.5)



```


```{r effect of Trial}

fulldata %>%     
   ggplot(aes(x = Genotype, y = Distance_travelled)) +
   geom_jitter(aes(shape = Genotype),
                size = 4) +
    geom_boxplot(aes(fill = Trial),
                 alpha = 0.25,
                 outlier.shape = NA
    ) +
   scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
    scale_y_continuous(labels = comma, name = "Distance (mm)") +
    ggtitle("Total distance moved by larvae in one hour per Trial (mm)") +
    facet_wrap(~Trial)



```

Overall distance travelled statistical analysis Genotype = 0.17, trial = 0.02
```{r Statistical analysis overall distance}

model_distance <- totaldist.hgsnat.5 %>%
  group_by(fish_id) %>%
  ungroup() %>%
  unique() %>%
  droplevels() %>%
  mutate(gTrial = interaction(Genotype, Trial, drop = TRUE)) %>%
  lmer(formula = Total_distance ~ Genotype + Trial + (1|gTrial))



plot(residuals(model_distance) ~ fitted(model_distance)) 
qqPlot(residuals(model_distance)) 

Anova(model_distance) %>%
  as.data.frame() %>%
  dplyr::rename(pval = `Pr(>Chisq)`) %>%
  kable(caption = "Type II Wald chisquare test of LME") %>%
  kable_styling()

```

em plots for overall distance travelled

```{r emmean distance}
print(emmeans(model_distance, ~ Genotype), type ="emmean") %>%
    as_tibble() %>%
    ggplot(aes(x = Genotype, y = emmean, colour = Genotype)) +
    geom_col(aes(fill = Genotype),
             alpha = 0.5,
             width = 0.75,
             position = position_dodge()) +
    geom_errorbar(aes(ymin = lower.CL, ymax =upper.CL),
                  width = 0.125,
                  size = 1,
                  position = position_dodge(width = 0.75)) +
    scale_color_viridis_d(end = 0.75, option = "plasma") +
    scale_fill_viridis_d(end = 0.75, option = "plasma") +
    labs(y = "model-predicted distance travelled",
         x = "Genotype") +
    theme(text = element_text(size = 20))
 



```


homs seemed to move less throughout all trials
```{r Stats per bin}

lm_bin <- hgsnat.5 %>%  
  mutate(gTrial = interaction(Genotype, Trial, drop = TRUE), 
         fish_id = as.character(fish_id)) %>% # these are to define interaction effects in the model
  lmer(formula = sqrt(Distance_travelled) ~ (Genotype*Bin) + Trial + (1|gTrial) + (1|fish_id), 
       data = .)

plot(residuals(lm_bin) ~ fitted(lm_bin)) 
qqPlot(residuals(lm_bin)) 


Anova(lm_bin) %>%
  as.data.frame() %>%
  dplyr::rename(pval = `Pr(>Chisq)`) %>%
  kable(caption = "Type II Wald chisquare test of LME") %>%
  kable_styling()


#---- distance travelled by bin.

print(emmeans(lm_bin, ~ Genotype * Bin), type = "response") %>%
  as_tibble() %>%
  mutate(binforvis = case_when(
    Bin == 1 ~ "Start-0:20:00",
    Bin == 2 ~ "0:20:00-0:40:00",
    Bin == 3 ~ "0:40:00-1:00:00",
    )) %>%
  ggplot(aes(x = binforvis, y = response, colour = Genotype)) +
  geom_col(aes(fill = Genotype),
           alpha = 0.5,
                 position = position_dodge()) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                size = 1,
                position = position_dodge()) +
  theme(axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45),
        legend.position = "bottom") +
  scale_color_viridis_d(end = 0.75, option = "plasma") +
  scale_fill_viridis_d(end = 0.75, option = "plasma") +
  facet_wrap(~Bin)
 



```


NEED TO DO THE STATS WITH POISONN
```{r Freq in centre}

hgsnat.5 %>%
  ggplot(aes(x = Genotype, y = Freq_in_centre)) +
    geom_jitter(aes(colour = Genotype,
                    shape = Trial),
                size = 4) +
    geom_boxplot(aes(fill = Genotype),
                 alpha = 0.25,
                 outlier.shape = NA
    ) +
    scale_fill_viridis_d(end = 0.75) +
    scale_colour_viridis_d(end = 0.75) +
    scale_y_continuous(labels = comma, name = "Number of crosses") +
    ggtitle("Number of crosses into the central zone") +
    facet_wrap (~Bin)

#---NO GOOD
  hgsnat.5 %>%
glmer(Freq_in_centre ~ Genotype + Trial + (1|fish_id),
data = .,
family = poisson(link = "log"))

check_overdispersion(glm_crosses) #hella overdispersed 



#negative binomial does not have assumption of equal dispersion so I will use that one instead 
glm_crosses <-
hgsnat.5 %>%
glmer.nb(Freq_in_centre ~ Genotype + Trial + Bin + (1|fish_id), 
data = .)

Anova(glm_crosses) %>%
  as.data.frame() %>%
  dplyr::rename(pval = `Pr(>Chisq)`) %>%
  kable(caption = "Type II Wald chisquare test of LME") %>%
  kable_styling()
```


```{r Freq in centre estimated mean plots}

#overall crosses into centre
print(emmeans(glm_crosses, ~ Genotype), type ="response") %>%
    as_tibble() %>%
    ggplot(aes(x = Genotype, y = response, colour = Genotype)) +
    geom_col(aes(fill = Genotype),
             alpha = 0.5,
             width = 0.75,
             position = position_dodge()) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax =asymp.UCL),
                  width = 0.125,
                  size = 1,
                  position = position_dodge(width = 0.75)) +
    scale_color_viridis_d(end = 0.75, option = "plasma") +
    scale_fill_viridis_d(end = 0.75, option = "plasma") +
    labs(y = "model-predicted total distance travelled",
         x = "Genotype") +
    theme(text = element_text(size = 20))
 
#---- Setting it up by bin 



print(emmeans(glm_crosses, ~ Genotype * Bin), type = "response") %>%
  as_tibble() %>%
  mutate(binforvis = case_when(
    Bin == 1 ~ "Start-0:20:00",
    Bin == 2 ~ "0:20:00-0:40:00",
    Bin == 3 ~ "0:40:00-1:00:00",
  
    )) %>%
  ggplot(aes(x = binforvis, y = response, colour = Genotype)) +
  geom_col(aes(fill = Genotype),
           alpha = 0.5,
                 position = position_dodge()) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                size = 1,
                position = position_dodge()) +
  theme(axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45),
        legend.position = "bottom") +
  scale_color_viridis_d(end = 0.75, option = "plasma") +
  scale_fill_viridis_d(end = 0.75, option = "plasma") +
  facet_wrap(~Bin)
 

```
SO homs consistently had lower n of crosses into centre compared to other genos. Lets look at how long the spent in the centre.

```{r time spent in centre plot}

hgsnat.5 %>%
 ggplot(aes(x = Genotype, y = Time_in_centre)) +
    geom_jitter(aes(colour = Genotype,
                    shape = Trial),
                size = 4) +
    geom_boxplot(aes(fill = Genotype),
                 alpha = 0.25,
                 outlier.shape = NA
    ) +
    scale_fill_viridis_d(end = 0.75, option = "plasma") +
    scale_colour_viridis_d(end = 0.75, option = "plasma") +
    scale_y_continuous(labels = comma, name = "Time in centre (s)") +
    ggtitle("Total time (in seconds) spent in the central zone") +
    facet_wrap (~Bin)



```


```{r centre time spent stats + emplot}

lm_time <- hgsnat.5 %>%  
  mutate(gTrial = interaction(Genotype, Trial, drop = TRUE), 
         fish_id = as.character(fish_id)) %>%
  lmer(formula = sqrt(Time_in_centre) ~ (Genotype*Bin) + Trial + (1|gTrial) + (1|fish_id), 
       data = .)


plot(residuals(lm_time) ~ fitted(lm_time)) 
qqPlot(residuals(lm_time)) 

#----  overall time spent in centre
print(emmeans(lm_time, ~ Genotype), type ="response") %>%
    as_tibble() %>%
    ggplot(aes(x = Genotype, y = response, colour = Genotype)) +
    geom_col(aes(fill = Genotype),
             alpha = 0.5,
             width = 0.75,
             position = position_dodge()) +
    geom_errorbar(aes(ymin = lower.CL, ymax =upper.CL),
                  width = 0.125,
                  size = 1,
                  position = position_dodge(width = 0.75)) +
    scale_color_viridis_d(end = 0.75, option = "plasma") +
    scale_fill_viridis_d(end = 0.75, option = "plasma") +
    labs(y = "model-predicted time in centre",
         x = "Genotype") +
    theme(text = element_text(size = 20))
 

#---- time spent in centre by bin.

print(emmeans(lm_time, ~ Genotype * Bin), type = "response") %>%
  as_tibble() %>%
  mutate(binforvis = case_when(
    Bin == 1 ~ "Start-0:20:00",
    Bin == 2 ~ "0:20:00-0:40:00",
    Bin == 3 ~ "0:40:00-1:00:00",
    )) %>%
  ggplot(aes(x = binforvis, y = response, colour = Genotype)) +
  geom_col(aes(fill = Genotype),
           alpha = 0.5,
                 position = position_dodge()) +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                size = 1,
                position = position_dodge()) +
  theme(axis.text.x = element_text(hjust = 1,
                               vjust = 1,
                               angle = 45),
        legend.position = "bottom") +
  scale_color_viridis_d(end = 0.75, option = "plasma") +
  scale_fill_viridis_d(end = 0.75, option = "plasma") +
  facet_wrap(~Bin)
 



```


```{r ggplate}

totaldist.hgsnat.5 %>%
plate_plot(
           value = Total_distance,
           label = Total_distance,
           plate_size = 24,
           position = Position,
           plate_type = "round",
           scale = 1.5,
           label_size = 1.5
           ) +
  facet_wrap(~Trial) +
  


```









