---
title: "14dpf_naglu_20_second_bins"
output: html_document
date: "2023-02-06"
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

NOTE: Omitted samples 73 - 120 as they were from a different pairmate. Reserach suggests that parent personality can affect larval behaviour.
Here I load my libraries

```{r}
library(tidyverse)
library(magrittr)
library(readxl)
library(dplyr)
# statistical analusis
library(lme4)
library(broom)
library(car)
library(performance)
library(emmeans)
library(MASS)
library(see)
# data visualstion
library(scales)
library(ggpubr)
library(ggeasy)
library(ggfortify)
library(ggbeeswarm)
library(kableExtra)


theme_set(theme_bw())
```


#joining distance with genotype
This part is joining the two files together - the meta data of genotype and the distance data output

```{r}
#stb stands for small timebin
data.stb <-
  read_excel("Statistics 20s time bin 14dpf naglu.xlsx")
data.stb <-
  subset(data.stb, select = -c(`...1`,`Independent Variable Independent Variable Treatment`))


data.stb <-
  data.stb %>%
  rename(
    "Trial" = `...2`,
    "Position" = `...3`,
    "Bin" = `...4`,
    "min dist" = "Distance moved center-point Minimum mm",
    "Distance_travelled" = "Distance moved center-point Total mm",
    "Mean_Velocity" = "Velocity center-point Mean mm/s",
    "Freq_in_centre" = "In zone Centre / center-point Frequency" ,
    "Time_in_centre" = "In zone Centre / center-point Cumulative Duration s"
  )%>%
       mutate(Trial = str_remove(Trial, pattern = "Trial     ") %>%
         as.factor())
```

Turning Bins into factors


At this point, there is an issue with the number of bins I have, so selecting all of them as levels is a bit tricky, her is karissa's solution. She made a new table consisting of the bin column, she then linked them to a sequential order of number

temp <- data.stb$Bin %>%
  unique %>%
  as_tibble %>%
  set_colnames("Bin") %>%
  mutate(Bin_number = rownames(.))

data.stb <-
temp %>%
  left_join(data.stb, by = ("Bin"))


Lachie suggested another method which will allow me to do it without making a new column as I was having trouble selecting all 180 characters as levels. 

!!!! I dont have to add in the values themselves - I can use the column name itself to do it !!! interestingly when i hover over the column when the data is presented on a new tab it says its a factor with 1 level but if I open it on the console it shows 180 levels. Pretty, pretty, pretty, pretty, pretty annoying. 

```{r}
data.stb <- 
  data.stb %>%
  mutate(Bin = factor(Bin, levels = unique(Bin)))


```



setting up the meta file and working it into the data.stb file 

```{r}
meta.stb <- read_excel("/Users/Lakshay/Desktop/Honours/Experimental data/14 dpf/14dpf naglu genotype.xlsx") %>%
  subset(select = -`...5`) %>%
  mutate(Genotype = factor(Genotype, levels = unique(Genotype)),
         Trial = as.factor(Trial),
         Position = as.factor(Position)
  )

fulldata.stb <-
  meta %>%
  left_join(data.stb, by = c("Trial", "Position")) %>%
  na.omit

```

#Velocity data 

lets get some graphs on the roll - gonna try making a function first like karissa showed me for genotypes:  

plotGenoProps <- function(input, xaxis) {
  ggplot(data = input) +
    geom_bar(aes(x = Genotype))
}

plotGenoProps(input = fulldata, xaxis = Genotype)

so, the function is laid out as function(x,y) where x and y represent the data that will be plugged into the function for each situation. next, you write out the formulas out as normal but instead of selecting data you substitute with x or y. 

plotDistance will come in handy for other work.
    
```{r}

plotDistance <- function(input, dependent, name_yaxis, title){
  ggplot(data = input, aes(x = Genotype, y = dependent)) +
    geom_jitter(aes(colour = Genotype,
                  shape = Trial),
              size = 4) +
  geom_boxplot(aes(fill = Genotype),
               alpha = 0.25,
               outlier.shape = NA

  ) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(labels = comma, name = "name_yaxis") +
  ggtitle("title")
  
}


plotVelocity <- 
  function(input, dependent, independent, name_yaxis, title){
  ggplot(data = input, aes(x = independent, y = dependent)) +
  geom_line(aes(group = Genotype))
  }

    
plotVelocity(input = fulldata.stb, dependent = fulldata.stb$Mean_Velocity, independent = fulldata.stb$Bin, name_yaxis = "Average velocity (mm/s) per 20 seconds", title = "Average velocity of 14 day post fertilisation naglu larvae at different time points") +
  geom_label(aes(label = fish_id), data = . 
             %>% dplyr::filter(`Mean_Velocity` > 15))



```
Alright, so the titles dont work, probs due to the scale_ stuff being specific to boxplots? also, need group in larger time intervals.

Also, sample 11's velocity is an obvious tracking error (also confirmed by checking video), so i'm going to omit it. 

```{r}
fulldata.stb <-
subset(fulldata.stb, Mean_Velocity < 15)

```




Trying 2 minutes + making bin numbers

#2 minute bins 
```{r}
twomin.stb <- fulldata.stb %>%
  dplyr::select(Bin) %>%
  unique %>%
  as_tibble %>%
  set_colnames("Bin") %>%
  mutate(Bin_number = rownames(.))

twomin.stb$Bin_number <- as.numeric(twomin.stb$Bin_number) 
  
    

#twomin.stb$range <-
  #cut(twomin.stb$Bin_number, breaks=(c(6,12,18,24,30))) - this method is still time consuming, imma try something different
  
twomin.stb$range <- 
  cut(twomin.stb$Bin_number, breaks=c(0:30*6))


```

IT WORKED! Connected time points of 2 mins to a range of 6. now I can join the fulldata to it by Bin and sneak in the range that way. This will hopefully allow me to average my velocities by trial, position and range number?

NOW, joining this new range with the original fulldata.stb 

```{r}

twomin.stb <- fulldata.stb %>%
  right_join(twomin.stb, by = "Bin")

```

now, lets try to average out the fishid, velocity and range. Angel helped with the next part.

Initially, i was having issues with the mutate function which does not delete rows - so it didnt know what to do with the mean for all time bins in a group. Instead, she used summarise because it can delete rows. idk why but it didnt work either so she used summarise_at instead.


```{r}

twomin.stb <- twomin.stb %>%
  dplyr::select(fish_id, range, Mean_Velocity, Trial, Genotype, ) %>%
  group_by(fish_id, range, Trial, Genotype) %>%
  summarise_at("Mean_Velocity", mean)




```

Ok, now I will plot out the average velocities in 2 min bins to the function I made earlier.


```{r}
  
plotVelocity(input = twomin.stb, dependent = twomin.stb$Mean_Velocity, independent = twomin.stb$range, name_yaxis = "Average velocity (mm/s) per 2 minutes", title = "Average velocity of 14 day post fertilisation naglu larvae at different time points")


```


Not a good looking graph, maybe i should change it to single points and/or by trial


```{r}
twomin.stb %>%
ggplot(aes(x = range, y = twomin.stb$Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype
                  ),
                alpha = 0.5,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  facet_wrap(~Trial) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(name = "Velocity mm/s") +
    scale_x_discrete (name = "Time (in 2 minute time bins)") +
  ggtitle("Velocity per 2 minute time bin over five one-hour trials") +




#+
  #ggarrange(nrow = 1)
    
    ```

Ok, interesting stuff, seems like trial 1 and 4 have some strange inconsistencies with wild type but not hom or het. apart from that, the genotypes seem very similar, gonna try 20 second again but just a single trial. 


  
  
#20 second per trial

```{r}
small_t1 <- 
fulldata.stb %>%
  subset(Trial == 1) 


temp <-
small_t1 %>%
  dplyr::select(Bin) %>%
  unique %>%
  as_tibble %>%
  set_colnames("Bin") %>%
  mutate(Bin_number = rownames(.),
         Bin_number = as.numeric(Bin_number)) 

small_t1 <-
  small_t1 %>%
  left_join(temp, by = "Bin")


small_t1 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.25,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(labels = comma, name = "Velocity mm/s") +
  scale_x_continuous(name = "Velocity per 20 second time bin") +
  ggtitle("Trial 1")
 
ggsave(filename = "Trial 1 14dpf 20 second velocity intervals.png", width = 20, height = 15, units = "cm", dpi = 600, scale = 1)

  
```

#20 second per trial, Trial 2 
```{r}
small_t2 <- 
fulldata.stb %>%
  subset(Trial == 2) 

small_t2 <-
  small_t2 %>%
  left_join(temp, by = "Bin")


small_t2 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.5,
              size = 2) +
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(labels = comma, name = "Velocity mm/s") +
  scale_x_continuous(name = "Velocity per 20 second time bin") +
  ggtitle("Trial 2")

  ggsave(filename = "Trial 2 14dpf 20 second velocity intervals.png", width = 20, height = 15, units = "cm", dpi = 600, scale = 1)

```

#20 second per trial, Trial 3

```{r}
small_t3 <- 
fulldata.stb %>%
  subset(Trial == 3) 

small_t3 <-
  small_t3 %>%
  left_join(temp, by = "Bin")

small_t3 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.5,
              size = 2) +
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(labels = comma, name = "Velocity mm/s") +
    scale_x_continuous(name = "Velocity per 20 second time bin") +
  ggtitle("Trial 3")

  ggsave(filename = "Trial 3 14dpf 20 second velocity intervals.png", width = 20, height = 15, units = "cm", dpi = 600, scale = 1)

  
```
  homs have relatively consistent velocity, wt and hets show "waves" of activity - suggesting that the homs are active for a greater period of itme. 
  both het and homs have dencrease in velocity at the end of the trial, opposite for the wt.
  
  
  
```{r}
small_t4 <- 
fulldata.stb %>%
  subset(Trial == 4) 


small_t4 <-
  small_t4 %>%
  left_join(temp, by = "Bin")

small_t4 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.5,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(labels = comma, name = "Velocity mm/s") +
  scale_x_continuous(name = "Velocity per 20 second time bin") +
  ggtitle("Trial 4 20 second velocity intervals")

 ggsave(filename = "Trial 4 14dpf 20 second velocity intervals.png", width = 20, height = 15, units = "cm", dpi = 600, scale = 1)
  
```
  
  Similar to trial 1, trial 4 has significant difference in velocity of wt compared to homs and hets. potentially due to drilling noises. this suggests that 
  
  
```{r}
small_t5 <- 
fulldata.stb %>%
  subset(Trial == 5) 


small_t5 <-
  small_t5 %>%
  left_join(temp, by = "Bin")

small_t5 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.5,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  scale_y_continuous(labels = comma, name = "Velocity mm/s") +
  scale_x_continuous(name = "Velocity per 20 second time bin") +
  ggtitle("Trial 5 20 second velocity intervals") +
  theme(text = element_text(size = 20))           

 ggsave(filename = "Trial 5 14dpf 20 second velocity intervals.png", width = 20, height = 15, units = "cm", dpi = 600, scale = 1)
  
```  

MAKING IT INTO A GG ARRANGE - next time just facet wrap by trial.

```{r}
ggarrange(
           small_t1 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.25,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
   theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  ggtitle("Trial 1") +
  theme(legend.position = "none"),

           
         small_t2 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.25,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  ggtitle("Trial 2") +
  theme(legend.position = "none"),
           
           small_t3 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.25,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  ggtitle("Trial 3")+
  theme(legend.position = "none"),

small_t4 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.25,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  ggtitle("Trial 4") +
  theme(legend.position = "none"),

small_t5 %>%
ggplot(aes(x = Bin_number, y = Mean_Velocity)) +
    geom_jitter(aes(colour = Genotype),
                alpha = 0.25,
              size = 2) +
  
  geom_smooth(aes(group = Genotype,
                  colour = Genotype)) +
  scale_fill_viridis_d(end = 0.75) +
  scale_colour_viridis_d(end = 0.75) +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  ggtitle("Trial 5"),

nrow = 1


)
 
ggsave(filename = "ALL TRIALS 14dpf 20 second velocity intervals.png", width = 90, height = 15, units = "cm", dpi = 600, scale = 1)

 
```

  
  
  
  
